---
title: "Accessibility Calculations"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: pdf_document
---

**NOTE:** Notebook takes about 10 minutes to run.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we calculate the amenities at the level of pedshed to get cumulative accessibility scores.

# Preliminaries

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3)
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = "amenities_points.rda")
load(file = "urban_hammer_da.rda")
data(urban_types)
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```


Intersect amenities with walksheds to obtain the cumulative opportunities:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

amenities_walksheds_da <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_da)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Validation walksheds:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

amenities_walksheds_validation <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_validation)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Save amenities by walkshed:
```{r}
save(amenities_walksheds_da, file = "amenities_walkshed_da.rda", compress = "xz")
save(amenities_walksheds_validation, file = "amenities_walkshed_validation.rda", compress = "xz")
```

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds_da |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

Map accessibility by category of amenity:
```{r}
urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE)) |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Find the most 15-minutes DA:
```{r}
amenities_walksheds_summary |>
  filter(Financial == max(Financial) |
         Sustenance == max(Sustenance) |
         Healthcare == max(Healthcare) |
         Library == max(Library))
```

Looks like GeoUID == "35250420" has a really nice combination: max in banking, max in sustenance, pretty high in health, and one library.

How many DAs have zero amenities?
```{r}
zero_amenity_walksheds <- urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  group_by(GeoUID, Type) |>
  summarize(Number_of_Amenities = sum(Number_of_Amenities),
            .groups = "drop") |>
  filter(Number_of_Amenities == 0)
```

```{r}
zero_amenity_walksheds |>
  pull(Type) |>
  table()
```

Check to see if there are any DAs in this list that have both an Urban and a Suburban part:
```{r}
zero_amenity_walksheds |>
  group_by(GeoUID, Type) |>
  summarize(n = n()) |>
  filter(n > 1)
```

Nope, there are none.

What is the area of DAs with zero amenities? What is their population?
```{r}
zero_amenity_walksheds_summary <- zero_amenity_walksheds |>
              st_drop_geometry() |>
  left_join(data_da_2016,
            by = c("GeoUID")) |>
  rename(Type.1 = Type.x,
         Type = Type.y) |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(`Shape Area`),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population), 
         Prop_Tot_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "Zero Amenities")

zero_amenity_walksheds_summary
```

Find all minimal 15-minutes DAs, meaning those that have at least one of each amenity:
```{r}
minimal_amenities_walksheds <- amenities_walksheds_summary |>
  filter(Financial >= 1,
         Sustenance >= 1,
         Healthcare >= 1,
         Library >= 1)
```

How many minimal 15 min DAs are urban/suburban:
```{r}
minimal_amenities_walksheds |>
  pull(Type) |>
  table()
```

What is the area of DAs with minimal 15-min walk sheds? What is their population?
```{r}
minimal_amenities_walksheds_summary <- minimal_amenities_walksheds |>
              st_drop_geometry() |>
  left_join(data_da_2016,
            by = c("GeoUID")) |>
  rename(Type.1 = Type.x,
         Type = Type.y) |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(`Shape Area`),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population), 
         Prop_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "Minimal Amenities")

minimal_amenities_walksheds_summary
```

Populations by urban fabric type:
```{r}
city_wide_summary <- urban_hammer_da |>
  st_drop_geometry() |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(Shape.Area),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population),
         Prop_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "City Wide")

city_wide_summary
```

Proportion of suburban and urban population in zero amenity sheds:
```{r}
zero_amenity_walksheds_summary$Population/city_wide_summary$Population
```

About 50% of suburban population (which represent about 22% of the total population in Hamilton) live in DAs with zero amenities within the 15-min walk shed.

About 15% of urban population (representing about 3.7% of the total population of the city) live in DAs with zero amenities within the 15-min walk shed.

Proportion of suburban and urban population in minimal 15-min sheds:
```{r}
minimal_amenities_walksheds_summary$Population/city_wide_summary$Population
```

Barely 2.3% of suburban residents live in DAs with minimal 15-min walk sheds. They represent approximately 1% of the total population of the city.

More than 15% of urban population live in DAs with minimal 15-min walk sheds, and account for approximately 3.8% of the total population of the city. 

Compare amenities by category and urban fabric type:
```{r}
amenities_walksheds_summary |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE)) |>
  group_by(Amenity_Category,
           Type) |>
  skim()
```

Sample two walksheds with at least one amenity, one from urban fabric type and one suburban, and plot:
```{r}
seed <- sample.int(1000000, 1)
set.seed(seed)
seed

# 609054, 144675, 741580

sample_walksheds <- amenities_walksheds_da |>
  group_by(Type) |>
  slice_sample(n = 1)

ggplot()  + 
  # geom_sf(data = urban_types |>
  #               filter(Type != "Rural"),
  #         aes(fill = Type)) +
  geom_sf(data = walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          aes(fill = Type),
          color = "black")+
  geom_sf(data = amenities_walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          #aes(color = Category)
          )
```

Distance to amenities:
```{r}
load("ttm_walk_da.rda")
```


```{r}
amenities_by_h3 <- amenities_points |>
  st_drop_geometry() |>
  group_by(h3_index) |>
  summarize(n_amenities = n(),
            .groups = "drop")
```


Join the amenities to the travel time matrix by means of the h3 indices:
```{r}
ttm_walk_to_amenities <- ttm_walk_da |>
  left_join(amenities_by_h3,
            by = "h3_index") |>
  drop_na(n_amenities)
```

Also join the urban fabric type:
```{r}
ttm_walk_to_amenities <- ttm_walk_to_amenities |>
  left_join(urban_hammer_da |>
              st_drop_geometry() |>
              transmute(GeoUID, Type = Type.1),
            by = "GeoUID")
```

Distribution of travel times by fabric type:
```{r}
ttm_walk_to_amenities |> 
  ggplot() +
  geom_histogram(aes(x = travel_time, fill = Type),
                 position = "dodge")
```

Summarize travel times to amenities from DA centroids:
```{r}
junk <- ttm_walk_to_amenities |>
  select(GeoUID, Type, travel_time) |>
  group_by(GeoUID, Type) |>
  summarize(min_travel_time = min(travel_time),
            mean_travel_time = mean(travel_time),
            .groups = "drop")
```

Distribution of mean travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = mean_travel_time,
             color = Type)) +
  geom_density()
```

Distribution of min travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = min_travel_time,
             color = Type)) +
  geom_density()
```

See Millward, Spinney, and Scott (2013) for some distance-decay functions for active transport. The decay curves are very steep, so the difference in minimum time and mean time to destinations by urban fabric can potentially have a large impact on the decision to walk.

For all destinations this is the equation:
$$
\log(Y) = 1.282 - 0.048 \cdot t
$$
For all destinations this is the equation:
$$
\log(Y) = 1.035 - 0.032 \cdot t
$$

$Y$ is the percentage of trips and $t$ is the time in minutes.

```{r}
ggplot() +
  geom_function(fun = function(x) exp(1.282)*exp(-0.048)^x,
                color = "blue") +
  geom_function(fun = function(x) exp(1.035) * exp(-0.032)^x,
                color = "red") +
  #geom_vline(xintercept =  4.37) +
  #geom_vline(xintercept = 6.02) +
  xlim(1, 15)
```

The median of the minimum distance to any amenity in urban walksheds is 4.00 and in suburban sheds is 6.00. If we use the impedance function for all destinations, this means that the probability of taking an active trip is about 40\% lower in suburban sheds:
```{r}
exp(1.282)*exp(-0.048)^6.0/exp(1.282)*exp(-0.048)^4.0
```

