---
title: "Profiling Walksheds"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we profile the ped sheds based on their network attributes.

# Preliminaries

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:
```{r}
load(file = "amenities_walkshed.rda")
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
data("urban_types")
load(file = "walksheds.rda")
load(file = "walksheds_net_attributes.rda")
```

## Classification tree

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:
```{r}
d <- walksheds_net_attributes |>
                  drop_na()
```


We use a technique with evolutionary trees - an alternative to classification trees that is less affecte by the greediness of the algorithm. Since evolutionary trees have some randomness (for the initial values), we can try several to find whether there are patterns. In this case we try four different random seed. 

Tree 1 with seed 1 (912571):
```{r}
#seed1 <- sample.int(1000000, 1)
seed1 <- 912571
set.seed(seed1)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree1)
```

Tree 2 with seed 2 (94766):
```{r}
#seed2 <- sample.int(1000000, 1)
seed2 <- 94766
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree2)
```

Tree 3 with seed 3 (46125):
```{r}
#seed3 <- sample.int(1000000, 1)
seed3 <- 46125
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree3)
```

Tree 4 with seed 4 (556090)
```{r}
#seed4 <- sample.int(1000000, 1)
seed4 <- 556090
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree4)
```

```{r}
#seed4 <- sample.int(1000000, 1)
seed4 <- 556090
set.seed(seed4)
tree5 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree5)
```

Create a function to estimate the missclassification rate and the evaluation function:
```{r}
mc <- function(obj) 1 - mean(predict(obj) == d$Type)


evalfun <- function(obj) 2 * nrow(d) * mc(obj) + width(obj) * log(nrow(d))

#trees <- list("evtree" = tree1, "rpart" = rp, "ctree" = ct, "rpart2" = rp2, "ctree2" = ct2)
trees <- list("evtree 1" = tree1, "evtree 2" = tree2, "evtree 3" = tree3, "evtree 4" = tree4)

round(sapply(trees, function(obj) c("misclassification" = mc(obj), "evaluation function" = evalfun(obj))), digits = 7)

```

All trees have similar fit.

Plot the trees:
```{r}
tree1_plot <- ggparty(tree1) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree2_plot <- ggparty(tree2) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree3_plot <- ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree4_plot <- ggparty(tree4) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

(tree1_plot + tree2_plot) / (tree3_plot + tree4_plot)
```

Notice that transitivity and motifs are the dominant attributes in the classification trees. Trees 1, 3, and 4 have the same general structure, and in fact trees 3 and 4 are identical. Based on this we choose to use `tree3`.

Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```

Plot the tree with the rounded break labels:
```{r}
ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = str_replace(breaks_label, 
                                          "0.0207311636201516", 
                                          "0.021"))) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.0,
                     ids = c(1, 2, 5)) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.02,
                     ids = c(3, 4, 6, 7)) +
  geom_node_plot(gglist = list(geom_bar(aes(x = "", 
                                            fill = Type),
                                        position = position_fill()),
                               xlab("Type"),
                               theme_minimal()),
                 # draw only one label for each axis
                 shared_axis_labels = TRUE,
                 # draw line between tree and legend
                 legend_separator = FALSE
                 )
```

Find examples of ped sheds in each of the leafs and plot the network for illustration.

Leaf 1:
```{r}

seed <- sample.int(100000, 1)
set.seed(seed)
seed
#69717

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity < 0.0207311636201516,
                             motifs < 378, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 2 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#84451, 44785

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs >= 378, 
                             transitivity < 0.0207311636201516,
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 3 (suburban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#81461, 84451

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs < 1024, 
                             transitivity >= 0.0207311636201516,
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 4 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#52534

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs >= 1024, 
                             transitivity >= 0.0207311636201516,
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Add leaf labels to data:
```{r}
d <- d |>
  mutate(leaf = case_when(transitivity < 0.0207311636201516 & motifs < 378 ~ "Leaf 1",
                          transitivity < 0.0207311636201516 & motifs >= 378 ~ "Leaf 2",
                          transitivity >= 0.0207311636201516 & motifs < 1024 ~ "Leaf 3",
                          transitivity >= 0.0207311636201516 & motifs >= 1024 ~ "Leaf 4"),
         leaf = factor(leaf))
```

```{r}
skim(d)
```

```{r}
d |>
  left_join(walksheds |>
              select(-Type) |>
              mutate(area = st_area(geometry),
                     area = units::set_units(area, km^2) |> 
                       drop_units(),
                     area = log(area)) |>
              st_drop_geometry(),
            by = "GeoUID") |>
  lm(area ~ leaf, data = _) |>
  summary()
```

But what about accessibility??

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```


To map accessibility by category of amenity we first categorize the number of amenities:
```{r}
amentities_da <-  urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE))
```


Plot:
```{r}
amentities_da |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Pivot wider:
```{r}
junk <- amentities_da |>
  pivot_wider(names_from = "Amenity_Category",
              values_from = "Number_of_Amenities")
```


```{r}
junk2 <- d |>
  left_join(junk |>
              #select(-Type) |>
              st_drop_geometry(),
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~forcats::fct_drop(.x)))
```

```{r}
summary(junk2 |> select(Sustenance:Library))
```


## Clusters: try self-organizing maps (SOM)

```{r}
wsn.som <- trainSOM(x.data = walksheds_net_attributes |>
                       select(-c(GeoUID, Type)),
                     dimension = c(5,5), 
                     verbose = TRUE, 
                     nb.save = 5, 
                     topo = "hexagonal")
```

```{r}
wsn.som
```


```{r}
plot(wsn.som, what="energy")
```


```{r}
wsn.som$clustering
```

```{r}
wsn.som$clustering |> table()
```

```{r}
plot(wsn.som, what="obs", type="hitmap")
```

```{r}
summary(wsn.som)
```

```{r}
par(mfrow = c(2,2))
plot(wsn.som, what = "obs", type = "color", variable = 1)
plot(wsn.som, what = "obs", type = "color", variable = 2)
```

```{r}
plot(wsn.som, what = "add", type = "pie", variable = walksheds_net_attributes$Type) +
  scale_fill_brewer(type = "qual") + 
  guides(fill = guide_legend(title = "Type"))
```


```{r}
quality(wsn.som)
```


```{r}
plot(superClass(wsn.som))
```


```{r}
wsn.som <- trainSOM(x.data = walksheds_net_attributes |>
                       select(-c(GeoUID, Type)),
                     dimension = c(4, 4), #2,3
                     verbose = TRUE, 
                     nb.save = 5,
                     topo = "square")
```

```{r}
plot(wsn.som, what="energy")
```

```{r}
quality(wsn.som)
```





