---
title: "Which fifteen-minutes neighborhoods are dead-ends? An analysis of the network attributes of fifteen-minute pedsheds"
author:
  - name: Author One
    email: a1@example.com
    affiliation: University One
    correspondingauthor: true
    footnote: 1
  - name: Author Two
    email: a2@example.com
    affiliation: University Two
  - name: Author Three
    email: a3@example.com
    affiliation: University One
    footnote: 2
  - name: Author Four
    email: a4@example.com
    affiliation: University One
    footnote: 2
address:
  - code: University One
    organization: Department
    addressline: 1 main street
    city: City
    state: State
    postcode: 123456
    country: Country
  - code: University Two
    organization: Department
    addressline: 2 main street
    postcode: 2054
    city: City
    country: Country
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Fifteen-minutes neihborhoods, a form of normative chronourbanism based on cumulative opportunities, has aroused interest as a way to reduce the need for motorized travel, and increase the livability, convenience, and health of the public. At the core of this concept is a pedshed, an area defined by the walkable isochrone of the eponymous fifteen minutes. As the idea of fifteen-minutes neighborhoods develops traction in policy and planning circles, it seems timely to revisit the way street network design can support—or obstruct—the stated goal of preserving or creating walkable neighborhoods with essential amenities. In this paper we examine a sample ($n=834$) of fifteen-minutes pedsheds in Hamilton, a medium-sized city in Canada, and how their sizes relate to the attributes of the transportation network. The analysis reveals that network design in suburban Hamilton conspires against the creating of fifteen-minutes neighborhoods. Much of urban Hamilton, in contrast, already has the characteristics of fifteen-minutes neighborhoods. The research points to elements of network design that can help to discriminate between candidate neighborhoods and dead-ends, and that can provide parameters for the design of new developments.

keywords: 
  - Fifteen-minutes neighborhoods
  - Pedshed
  - Walkability
  - Accessibility
  - Network analysis
journal: "Sustainable Cities and Society"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography:
  - "`r system('kpsewhich ../bibliography/bibliography.bib', intern=TRUE)`"
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

```{r chunk-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      out.width = "0.8\\linewidth")
```

```{r load-packages}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(ggspatial) # Spatial Data Framework for ggplot2
library(glue) # Interpreted String Literals
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
#library(gt) # Easily Create Presentation-Ready Display Tables
#library(rpart) # Recursive Partitioning and Regression Trees
#library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
#library(patchwork) # The Composer of Plots
#library(partykit) # A Toolkit for Recursive Partytioning
library(tidyr) # Tidy Messy Data
#library(tree) # Classification and Regression Trees
library(scales) # Scale Functions for Visualization
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
#library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r load-data}
 # From package vaccHamilton
data("urban_types")

# Load trained evolutionary trees
load(file = glue::glue(here::here(),
                       "/data/trees.rda"))

# Analysis-ready data. See notebooks in folder data-raw for details
load(file = glue::glue(here::here(),
                       "/data/amenities_walkshed_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/hamilton_net.rda"))
load(file = glue::glue(here::here(),
                       "/data/urban_hammer_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_da.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_da_net_vars.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_validation.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_validation_net_vars.rda"))
load(file = glue::glue(here::here(),
                       "/data/walksheds_motif_3_count.rda"))
```

```{r}
# Bind the centroid and non-centroid walksheds:
walksheds <- bind_rows(walksheds_da,
                       walksheds_validation) |>
  mutate(origin_type = rep(c("centroid", "non-centroid"), each = nrow(walksheds_da)))
```

```{r }
# Obtain training and validation datasets by sampling from centroid and non-centroid walksheds:

# Set seed for replicability
seed <- 3626425
set.seed(seed)

# slice_sample() will select one of two rows for each DA at random
walksheds_train <- walksheds |>
  group_by(GeoUID) |>
  slice_sample(prop = 0.5)

# The antijoin will keep the rows in walksheds that do not appear in walksheds_train
walksheds_test <- walksheds |>
  anti_join(walksheds_train |>
              st_drop_geometry() |>
              select(GeoUID, origin_type),
            by = c("GeoUID", "origin_type"))
```

```{r bind-centroid-non-centroid-network-variables}
# Bind the centroid and non-centroid dataframes with walkshed network variables:
walksheds_net_vars <- bind_rows(walksheds_da_net_vars,
                                walksheds_validation_net_vars) |>
  mutate(origin_type = rep(c("centroid",
                             "non-centroid"), 
                           each = nrow(walksheds_da)))
```

```{r obtain-training-testing-network-variables}
# Now filter from the network variables to obtain training and test datasets:
walksheds_train_net_vars <- walksheds_net_vars |>
  semi_join(walksheds_train |>
              st_drop_geometry() |>
              select(GeoUID, origin_type),
            by = c("GeoUID", "origin_type"))

walksheds_test_net_vars <- walksheds_net_vars |>
  semi_join(walksheds_test |>
              st_drop_geometry() |>
              select(GeoUID, origin_type),
            by = c("GeoUID", "origin_type"))
```

```{r training-and-testing-dataframes-for-evolutionary-tree}
d <- walksheds_train_net_vars |>
  select(-girth) |>
  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)

d_test <- walksheds_test_net_vars |>
  select(-girth) |>
  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)

d_all <- bind_rows(d, 
                   d_test)
```

# Introduction

> "Il faut oublier la traversée de Paris d'est en ouest en voiture"^[ "We must forget about crossing Paris from east to west by car"]. 
> --Anne Hidalgo

With this declaration during her reelection campaign in 2020, Anne Hidalgo brought international attention to the concept of walkable fifteen-minutes neighborhoods [@alimi2020]. 


<!-- 
Automotive retrofits in traditional city centers:

@bass2019automotive
-->

@knight2018walkable
@liu2022toward
@pozoukidou2021fifteen
@weng2019fifteen

We address the following research questions:

1. Is there a difference in the transportation network topology between urban and suburban areas within 15-minute walking distances?
2. If such a difference exists, what is its effect on the 15-minute walksheds in terms of accessibility, where accessibility is understood as the provision of opportunities?

# Data

Relevant data to develop this paper comprises to use 15-min pedestrian sheds (or walksehds) classified into urban and suburban types in Hamilton.

15-minute walksheds in the city of Hamilton were obtained by extracting relevant structural attributes of the street network within 15-minute pedestrian sheds. These sheds were calculated for each shed as polygon representations of origin-destination walking time matrices, with a maximum travel time of 15 minutes. Walkshed calculations were generated using the travel_time_matrix function for walking from the {r5r} R package. This process involved defining the street network, as well as the origins and destinations of walking trips. The street network data for these calculations was sourced from OpenStreetMap @OpenStreetMap2023.   
<!--- 
To see the details of the routing see notebook 01-OSM-Network-and-Routing.Rmd in folder data-raw
--->

The origins we established are proxies for the centroids of neighborhoods in the urban and peri-urban areas of the city of Hamilton. Specifically, these are the centroids of Hamilton's Dissemination Areas (DA). A DA is composed of one or more neighboring dissemination blocks and represents the smallest standard geographic area for which all census data are disseminated in Canada. Data on urban and sub-urban boundaries, DAs,and population data in 2016 were retrieved from @HamiltonData2023.

Destinations were established as the centroids of geohashed hexagons within the region of interest, where Hamilton's amenities are located, assuming that presence of amenities is a proxy to provision of opportunities. These destinations were chosen to avoid calculating travel times to each individual amenity. Instead, we used the geohashed polygons as destinations, identifying the amenities within each polygon and assuming that the centroid of each hexagon represents the location of the amenities within it. Specifically, we retrieved geohashing data for Hamilton using the H3 system @h3Package and generated child hexagons at resolution 13, with an average area of 43.87 $m^2$. Amenities location and their typology were downloaded from @OpenStreetMap2023. 
<!--- 
To see the details of the destinations see notebook 02-OSM-Amenities.Rmd in folder data-raw
--->


# Methods

The proposed method involves studying 15-minute walking neighborhoods in two stages. In the first stage, we assess the accessibility of Hamilton's walksheds from a normative perspective, focusing on the availability of opportunities. In the second stage, we analyze the topology of urban and suburban walksheds based on various network structure attributes. Finally, we examine whether there are differences in transportation network topology between urban and suburban areas, and analyze how these differences affect access to opportunities within walksheds.

## Calculation of accessibility levels

The indicator we use assesses levels of accessibility from a normative perspective, focusing on reachable opportunities as estimates of the potential for spatial interactions @Hansen1959. The advantage of using this form of accessibility analysis in this study is that it has been shown to be a valuable method for exploring the relationship between transportation infrastructure and urban structure @Handy1997.

In this study, we calculate cumulative accessibility scores by summing the number of amenities within urban and suburban walksheds, categorized into the four types provided by the data source: financial, sustenance, healthcare, and library. Additionally, we extract population data of each walkshed. Summary statistics of population are also included in the analysis.

Additionally, the accessibility assessment is completed by obtaining summary statistics of travel times in urban and suburban walksheds. To study the impact of urban and suburban walksheds on walking decisions, minimum and mean travel times to destinations are introduced into time-decay functions for walking decisions, as developed by Milward (2013). These are active travel impedance functions based on data from Canada.


<!--- 
To see the details of the accessibility calculations, summary statistics, and obtention of impedance functions see notebook 04-Accessibility-Calculations.Rmd in folder data-raw
--->

## Urban and suburban walkshed modelling

### Network attributes by walkshed

In this work we characterize urban and suburban walksheds based on their network structure. As a way to capture and describe the structure of networks we calculate a selection of 12 network structural attributes using the network analysis package {igraph} [@IgraphManual]. The set of attributes that we consider for this work is described Table \ref{tab:network-attributes}.
```{r network-attributes-descriptive-table}
data.frame(attribute = c("Number of vertices",
                   "Number of edges",
                   "Diameter",
                   "Radius",
                   "Global efficiency",
                   "Edge density",
                   "Mean distance",
                   "Girth",
                   "Minimum cut",
                   "Edge connectivity",
                   "Transitivity",
                   "Motifs (v)"),
           description = c("The total number of vertices in a graph.",
                           "Total number of edges in a graph.",
                           "The longest shortest path in a graph, akin to the longest segment of a line that cuts a circle.",
                           "Minimum of all the maximum distances between each vertex and all other vertices.",
                           "The average of the inverse distances between all pairs of vertices.",
                           "The ratio of the number of edges in a graph to the maximum number of possible edges (i.e., the number of edges of a point-to-point network.",
                           "Mean length of all shortest paths from each vertex.",
                           "The length of the shortest cycle, where a cycle is a trail where only the starting and ending point are a repeated vertex.",
                           "The smallest number of edges that, when removed, separate a graph into two or more components.",
                           "Minimum number of nodes that, when removed, disconnect two nodes.",
                           "Probability that a vertex neighbors are neighbors between them.",
                           "Isomorphic subgraphs when the number of nodes is (v). A graph with few motifs will have many unique elements; a graph with a large number of motifs will have a small number of patterns that are repeated. In this analysis we calculate motifs of size v=3 and v=4.")) |>
  kableExtra::kable(booktabs = TRUE,
                    caption = "\\label{tab:network-attributes}Network attributes in the analysis") |>
  kableExtra::kable_styling(font_size = 7, latex_options = "striped") |>
  kableExtra::column_spec(2, width = "40em")
```

<!--- 
To see the details of the attribute calculations see notebook 03-Network-Attributes-by-Walkshed.Rmd in folder data-raw
--->

### Modelling

To model urban and suburban walksheds, we employ a classification tree approach using the set topology attributes calculated in the previous stage. Classification trees are machine learning techniques used for both regression and classification, widely appreciated in various disciplines for their ability to represent knowledge through decision rules, making them easy to interpret and analyze @Loh2011 @James2013 @Paez2019Trees. Specifically, we utilize evolutionary trees, a variant of decision trees that are less constrained by the limitations of greedy top-down induction algorithms @Cha2009 @Kretowski2007 @Grubinger2014. Since evolutionary trees use a genetic algorithm that introduces randomness in the initial conditions, we run the model from the R package evtree @Grubinger2014 with 20 random seeds to identify the network variables that consistently characterize urban and suburban walksheds.

Finally we study effects in accessibility

<!--- 
To see the details of walkshed profile calculations see notebook 06-Walkshed-Profiles.Rmd in folder data-raw
--->

## Walkshed size in urban and suburban areas




# Results and discussion

## Classification of walksheds based on network attributes

```{r best-tree}
# Retrieve the best evolutionary tree. The algorithm of evolutionary trees has some randomness due to random initial values. We compared a number of trees and chose the one that gave the best fit in terms of missclassification and size of the model. See notebook `05-Walkshed-Profiles.Rmd in folder `data-raw` for details.

best_evtree <- trees[[4]]
```

```{r generate-plot-best-tree}
# Round the labels of the edges. The relevant information is in a dataframe produced by ggparty()
ggparty_plot <- best_evtree |>
  ggparty()

# Round the number of digits and keep the inequality signs
ggparty_data <- ggparty_plot$data |>
  mutate(breaks_label = paste0(substr(breaks_label, 4, 5), 
                               substr(breaks_label,10, 17)),
         splitvar = case_when(splitvar == "edge_density" ~ "Edge Density",
                              splitvar == "transitivity" ~ "Transitivity",
                              splitvar == "normalized_motifs_3" ~ "Normalized Motifs (3)",
                              is.na(splitvar) ~ NA))

# Replace the dataframe in the ggplot object
ggparty_plot$data <- ggparty_data 

evtree_plot <- ggparty_plot +
  geom_edge() +
  geom_node_splitvar(id = "inner", size = 3) +
  # parse should be set to FALSE so the labels do not try to parse the inequality signs
  geom_edge_label(mapping = aes(label = breaks_label),
                  parse = FALSE,
                  size = 2.5) +
  # # pass list to gglist containing all ggplot components we want to plot for each
  # # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = Type),
                                        position = position_fill()),
                               xlab(""),
                               ylab(""),
                               scale_fill_manual(name = "Urban fabric type",
                                                 values = c("#FDAE61", "#1A9850")),
                               theme(axis.text.y = element_blank(),
                                     panel.background = element_rect(fill = NA),
                                     axis.ticks = element_blank(),
                                     plot.background =  element_rect(fill = NA)))) +
  geom_node_label(line_list = list(aes(label = paste0("Leaf ", ifelse(id >= 6, id - 3, id - 2)))),
                  line_gpar = list(list(size = 7, parse = F)),
                  ids = "terminal",
                  nudge_x = 0.02,
                  nudge_y = -0.4)
```

```{r get-legend-for-tree}
tree_legend <- ggplot(d) +
  geom_bar(aes(x = Type, fill = Type)) +
  scale_fill_manual(name = "Urban fabric type",
                    values = c("#FDAE61", "#1A9850")) +
  #theme_void() +
  guides(fill = guide_legend(nrow = 1)) #+
  #theme(legend.position = "bottom")

lgnd <- cowplot::get_legend(tree_legend)
```


```{r plot-evtree, fig.cap="\\label{fig:evolutionary-tree}Evolutionary classification tree. The dependent variable is the fabric type: Suburban (orange), Urban (green)."}

evtree_plot <- cowplot::plot_grid(evtree_plot, lgnd,
                   nrow = 2,
                   rel_heights = c(1, 0.1))


cowplot::save_plot(evtree_plot,
                   nrow = 2,
                   base_width = 7,
                   filename = glue::glue(here::here(),"/plots/evtree.png"))

evtree_plot
```

```{r add-classification-to-train-data}
# Use the classification tree to add classes to dataframes

# Training dataset
d <- d |>
  mutate(leaf = case_when(edge_density < 0.00590599876314162 & normalized_motifs_3 < 0.865853658536585  ~ "Leaf 1",
                          edge_density < 0.00590599876314162 & normalized_motifs_3 >= 0.865853658536585 ~ "Leaf 2",
                          edge_density >= 0.00590599876314162 & normalized_motifs_3 < 0.967078189300412 ~ "Leaf 3",
                          edge_density >= 0.00590599876314162 & normalized_motifs_3 >= 0.967078189300412 ~ "Leaf 4") ,
         leaf = factor(leaf))

# All dataset (training and validation pooled)
d_all <- d_all |>
  mutate(leaf = case_when(edge_density < 0.00590599876314162 & normalized_motifs_3 < 0.865853658536585  ~ "Leaf 1",
                          edge_density < 0.00590599876314162 & normalized_motifs_3 >= 0.865853658536585 ~ "Leaf 2",
                          edge_density >= 0.00590599876314162 & normalized_motifs_3 < 0.967078189300412 ~ "Leaf 3",
                          edge_density >= 0.00590599876314162 & normalized_motifs_3 >= 0.967078189300412 ~ "Leaf 4") ,
         leaf = factor(leaf))
```

```{r}
# Compute the purity of the leafs as per training
purity <- d |> 
  group_by(leaf) |> 
  summarize(urban_fabric = sum(ifelse(Type == "Urban", 
                                      1, 
                                      0))/n(),
            .groups = "drop") |> 
  mutate(purity = ifelse(leaf == "Leaf 1" | leaf == "Leaf 3", 
                         1 - urban_fabric, 
                         urban_fabric))
```


Evolutionary trees: discuss training procedure. We used a training sample and a test sample. We trained n=30 trees per batch, and saved the results for the batch with the best tree (5 batches). The best tree was selected based on the evaluation function (write equation and explain).

The results of training an evolutionary classification tree are shown in Fig. \ref{fig:evolutionary-tree}. This model addresses our first research question (are networks different between urban and suburban regions?). The answer to this question is unambiguously affirmative. As seen in the figure, the tree has four terminal nodes (or leaves). The purity of the leaves (meaning the mix of classes in a leaf) is very high, since each leaf is composed largely of a single class (`r round(purity$purity[1], 3) * 100`\% and  `r round(purity$purity[3], 3) * 100`\% Suburban in Leaves 1 and 3 respectively, and `r round(purity$purity[2], 3) * 100`\% and `r round(purity$purity[4], 3) * 100`\% Urban in Leaves 2 and 4 respectively). Thus, there are two kinds each of urban and suburban walksheds, depending on the attributes of the network that generated them. 

Of the attributes in Table \ref{tab:network-attributes}, only two are picked up by the classification algorithm: edge density and motifs of size 3. The first attribute that helps to discriminate between urban and suburban walksheds is edge density. The cutoff value of edge density is 0.00590. Recall that edge density is the number of edges in the graph relative to the maximum possible number of edges--that is, compared to a fully connected graph. This means that the cutoff is about 0.59% of the number of edges in a fully connected graph with the same number of vertices. The range of values of this variable in our walksheds is `r round(d_all |> pull(edge_density) |> min(), 5) * 100`\% and `r round(d_all |> pull(edge_density) |> max(), 5) * 100`\% (the mean is `r round(d_all |> pull(edge_density) |> mean(), 5) * 100`\%). But there are marked differences in the value of this attribute by Leaf in the classification tree, as seen in Table \ref{tab:skim-of-edge-density}

```{r walkshed-edge-density-by-leaf}
skim_edge_density <- d_all |>
  group_by(leaf) |>
  skim_without_charts(edge_density) |>
  yank("numeric") |>
  mutate(leaf_desc = c("Suburban (Type I: low edge density)", "Urban (Type I: low edge density)",
                       "Suburban (Type II: high edge density)", "Urban (Type II: high edge density)")) |>
  select(-c(skim_variable, n_missing, complete_rate)) |>
  select(leaf, leaf_desc, everything())

skim_edge_density |>
  kable(booktabs = TRUE,
        digits = 5,
        col.names = c("leaf", "leaf description", "mean", "s.d.", "min", "first quartile", "median", "third quartile", "max"),
        caption = "\\label{tab:skim-of-edge-density}Descriptive statistics of edge density by Leaf in classification tree") |>
  kableExtra::kable_styling(font_size = 7, latex_options = "striped")
```

<!--
NOTE FOR DICUSSION: WHY IS THE EDGE DENSITY HIGHER IN SUBURBAN REGIONS?
-->

Table \ref{tab:skim-of-motifs} shows the values of the motifs of size 3 by Leaf. 

```{r walkshed-motifs-by-leaf}
d_all |>
  group_by(leaf) |>
  skim_without_charts(normalized_motifs_3) |>
  yank("numeric") |>
  mutate(leaf_desc = c("Suburban (Type I: low edge density)", "Urban (Type I: low edge density)",
                       "Suburban (Type II: high edge density)", "Urban (Type II: high edge density)")) |>
  select(-c(skim_variable, n_missing, complete_rate)) |>
  select(leaf, leaf_desc, everything())

skim_edge_density |>
  kable(booktabs = TRUE,
        digits = 5,
        col.names = c("leaf", "leaf description", "mean", "s.d.", "min", "first quartile", "median", "third quartile", "max"),
        caption = "\\label{tab:skim-of-motifs}Descriptive statistics of motifs of size 3 by Leaf in classification tree") |>
  kableExtra::kable_styling(font_size = 7, latex_options = "striped")
```

The differences between walksheds in various types of urban and suburban fabric are better appreciated by means of some examples. Fig. REF 

```{r}
walksheds <- walksheds |> 
  left_join(d_all |>
              st_drop_geometry() |>
              select(GeoUID,
                     origin_type,
                     leaf),
            by = c("GeoUID", 
                   "origin_type"))
```

```{r select-plot-leaf-1}
set.seed(29919)

walkshed_leaf_1 <- walksheds |> 
  #ungroup () |>
  filter(leaf == "Leaf 1") |>
  slice_sample(n = 1)

leaf_1_plot <- hamilton_net$edges |>
  st_intersection(walkshed_leaf_1)

leaf_1_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_1_plot$edge_index) |>
  st_transform(crs = st_crs(urban_types))

walkshed_leaf_1 <- walkshed_leaf_1 |> 
  st_transform(crs = st_crs(urban_types))

# Find the center of the bounding box; we'll use it to plot on a false origin and fixed extent
max_x_1 <- st_bbox(leaf_1_plot)[3] - st_bbox(leaf_1_plot)[1]
max_y_1 <- st_bbox(leaf_1_plot)[4] - st_bbox(leaf_1_plot)[2]
center_1 <- c(max_x_1/2 + st_bbox(leaf_1_plot)[1], max_y_1/2 + st_bbox(leaf_1_plot)[2])
```

```{r select-plot-leaf-2}
set.seed(29919)

walkshed_leaf_2 <- walksheds |> 
  filter(leaf == "Leaf 2") |>
  slice_sample(n = 1)

leaf_2_plot <- hamilton_net$edges |>
  st_intersection(walkshed_leaf_2)

leaf_2_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_2_plot$edge_index) |>
  st_transform(crs = st_crs(urban_types))

walkshed_leaf_2 <- walkshed_leaf_2 |> 
  st_transform(crs = st_crs(urban_types))

# Find the center of the bounding box; we'll use it to plot on a false origin and fixed extent
max_x_2 <- st_bbox(leaf_2_plot)[3] - st_bbox(leaf_2_plot)[1]
max_y_2 <- st_bbox(leaf_2_plot)[4] - st_bbox(leaf_2_plot)[2]
center_2 <- c(max_x_2/2 + st_bbox(leaf_2_plot)[1], max_y_2/2 + st_bbox(leaf_2_plot)[2])
```

```{r select-plot-leaf-3}
set.seed(29922)

walkshed_leaf_3 <- walksheds |> 
  filter(leaf == "Leaf 3") |>
  slice_sample(n = 1)

leaf_3_plot <- hamilton_net$edges |>
  st_intersection(walkshed_leaf_3)

leaf_3_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_3_plot$edge_index) |>
  st_transform(crs = st_crs(urban_types))

walkshed_leaf_3 <- walkshed_leaf_3 |> 
  st_transform(crs = st_crs(urban_types))

# Find the center of the bounding box; we'll use it to plot on a false origin and fixed extent
max_x_3 <- st_bbox(leaf_3_plot)[3] - st_bbox(leaf_3_plot)[1]
max_y_3 <- st_bbox(leaf_3_plot)[4] - st_bbox(leaf_3_plot)[2]
center_3 <- c(max_x_3/2 + st_bbox(leaf_3_plot)[1], max_y_3/2 + st_bbox(leaf_3_plot)[2])
```

```{r select-plot-leaf-4}
set.seed(299192245)

walkshed_leaf_4 <- walksheds |> 
  filter(leaf == "Leaf 4") |>
  slice_sample(n = 1)

leaf_4_plot <- hamilton_net$edges |>
  st_intersection(walkshed_leaf_4)

leaf_4_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_4_plot$edge_index) |>
  st_transform(crs = st_crs(urban_types))

walkshed_leaf_4 <- walkshed_leaf_4 |> 
  st_transform(crs = st_crs(urban_types))

# Find the center of the bounding box; we'll use it to plot on a false origin and fixed extent
max_x_4 <- st_bbox(leaf_4_plot)[3] - st_bbox(leaf_4_plot)[1]
max_y_4 <- st_bbox(leaf_4_plot)[4] - st_bbox(leaf_4_plot)[2]
center_4 <- c(max_x_4/2 + st_bbox(leaf_4_plot)[1], max_y_4/2 + st_bbox(leaf_4_plot)[2])
```

```{r max-extent}
# Find the max extent in x and y to ensure that all plots are at the same scale
max_x <- max(c(max_x_1, max_x_2, max_x_3, max_x_4))
max_y <- max(c(max_y_1, max_y_2, max_y_3, max_y_4))
```

```{r generate-plots-for-all-leaves}
leaf_1_gg <- ggplot() +
  geom_sf(data = walkshed_leaf_1 |>
            mutate(geometry = geometry - center_1),
          color = "#FDAE61",
          fill = "#FDAE61") +
  geom_sf(data = leaf_1_plot |>
            mutate(geometry = geometry - center_1)) +
  coord_sf(xlim = c(-max_x/2, max_x/2),
           ylim = c(- max_y/2, max_y/2)) +
  theme_void()

leaf_2_gg <- ggplot() +
  geom_sf(data = walkshed_leaf_2 |>
            mutate(geometry = geometry - center_2),
          color = "#1A9850",
          fill = "#1A9850") +
  geom_sf(data = leaf_2_plot |>
            mutate(geometry = geometry - center_2)) +
  coord_sf(xlim = c(-max_x/2, max_x/2),
           ylim = c(- max_y/2, max_y/2)) +
  theme_void()

leaf_3_gg <- ggplot() +
  geom_sf(data = walkshed_leaf_3 |>
            mutate(geometry = geometry - center_3),
          color = "#FDAE61",
          fill = "#FDAE61") +
  geom_sf(data = leaf_3_plot |>
            mutate(geometry = geometry - center_3)) +
  coord_sf(xlim = c(-max_x/2, max_x/2),
           ylim = c(- max_y/2, max_y/2)) +
  theme_void()

leaf_4_gg <- ggplot() +
  geom_sf(data = walkshed_leaf_4 |>
            mutate(geometry = geometry - center_4),
          color = "#1A9850",
          fill = "#1A9850") +
  geom_sf(data =leaf_4_plot |>
            mutate(geometry = geometry - center_4)) +
  coord_sf(xlim = c(-max_x/2, max_x/2),
           ylim = c(- max_y/2, max_y/2)) +
  theme_void()
```


```{r sample-sheds-by-leaf-plot, fig.cap="\\label{fig:sample-sheds}Examples of walksheds and their networks by Leaf in the classifactions."}
sample_sheds_by_urban_type <- cowplot::plot_grid(leaf_1_gg, leaf_2_gg, leaf_3_gg, leaf_4_gg,
                   nrow = 2, 
                   ncol = 2,
                   labels = c("Leaf 1 (Suburban Type I)", "Leaf 2 (Urban Type I)", 
                              "Leaf 3 (Suburban Type II)", "Leaf 4 (Urban Type II)"),
                   label_size = 7)

cowplot::save_plot(sample_sheds_by_urban_type,
                   filename = glue::glue(here::here(),"/plots/sample_sheds_by_urban_type.png"),
                   nrow = 2,
                   ncol = 3,
                   base_width = 3.5)

sample_sheds_by_urban_type
```


Where are these found?
```{r map-of-leaves, fig.cap="\\label{fig:leaves-by-DA}Distribution of Leaf classes in urban and suburban Hamilton (gray area is rural Hamilton)."}
ggplot() +
  geom_sf(data = hamilton_cma,
          fill = "lightgray") +
  geom_sf(data = urban_hammer_da |> 
  left_join(d |> 
              select(GeoUID, 
                     leaf), 
            by = "GeoUID"),
          aes(fill = leaf)) +
  scale_fill_manual(values = c("Leaf 1" = scales::muted("#FDAE61"), "Leaf 2" = scales::muted("#1A9850"), "Leaf 3" = "#FDAE61", "Leaf 4" = "#1A9800")) +
  coord_sf(xlim = c(-80.04, -79.63),
           ylim = c(43.115, 43.35)) +
  theme_void() +
  theme(legend.position = "bottom")
```

DISCUSS: there appear to be significant (?) differences between the sizes of the walksheds. 

## Area of 15-min walksheds

Walksheds in suburban areas tend to be smaller, walksheds in urban areas larger. Are there differences between walksheds of Type I and of Type II? Test differences of means.

Another feature of suburban environments is that they tend to be lower density. Proceed to accessibility calculations.

## 15-min accessibility analysis

Words go here.

## Implications for walking behavior

Use Millward 
# References {-}




